---
title: "Tutorial for Conditionals, Logic, and Programming Flow"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to use conditionals, logic, and programming flow in R!
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(learnr)
```

## Welcome

This is a demo tutorial. 

###

In this tutorial, you will learn how to:

  * Use `if`, `else`, and `for`
  * Use `&` and `|`
  * Know when to use `in` and `%in%`
  * Get comfortable using `==`, `!=`, and `!`

### `if` 

`if` let's you evaluate code dependent on some condition specified in R. 

For example, the following code will print "hello world!" on the console because the condition given is `TRUE`:

```{r example-if, echo=TRUE, eval=FALSE}
if (TRUE) { 
  print("hello world!") 
}
```

How could you modify the following code to print "Battlestar Galactica" to the console?

```{r if1, exercise=TRUE, exercise.eval=FALSE}
if (_____) {
  print("Battlestar Galactica")
}
```

```{r if1-solution}
if (TRUE) {
  print("Battlestar Galactica")
}
```

### Using Meaningful Conditions

The conditions given can get more complicated.  For example, what do you think the following code might do?

```{r if2, eval = FALSE, echo = TRUE}
x <- 1
if (x > 0) {
  print("Pokemon Go!")
}
```

It prints "Pokemon Go!"

### An Applied Example 

Sometimes when you're working with data, some functions you'll want to use may require data frames, while others require matrices.  In the case that you want your code to convert your data to a data frame before passing the data to another function, you might use code that looks like this: 

```{r, eval = FALSE, echo = TRUE}
if (! is.data.frame(df)) {
  df <- as.data.frame(df)
}
```

Of course, you have to be careful when converting your data to a data frame since it needs to be in an appropriate format to be converted to a data frame (like a matrix or table).  

### Multiple Branches 

A programmer can use `if` and `else` together to evaluate one of two programming expressions dependent on the condition given. 

For example, what do you think the following might do? 

```{r logicals, echo = FALSE}
question("What will the following code return? 
  ```{r, echo = TRUE, eval = FALSE}
  x <- 123
  if (is.character(x)) {
   print('The Simpsons')
  } else {
   print('Sesame Street')
  }
  ```
",
         answer("The Simpsons", message = "x isn't a character object, so this answer is incorrect."),
         answer("Sesame Street", correct = TRUE, message = "x is not a character object, so the code in the `else` block is what gets evaluated, resulting in 'Sesame Street' being printed to the console"),
         answer("Blues Clues", message = "This didn't even appear in any of the branches in the code, so there's no way this could have been the output."),
         allow_retry = TRUE
)
```